SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-12-05 related.tex
Last sub-section still needs work!

Perceptron-based prediction mechanisms have proven to yield superior accuracy, 
which represents an opportunity for optimizations in cache management. 
%Khan \textit{et. al.} proposed SDBP~\cite{sdbp}, in this work, a sampler 
%structure keeps partial tags of sampled sets separate from the cache. 
%Three tables of two-bit saturating counters are accessed using a technique 
%similar to a skewed branch predictor~\cite{Piece_Linear}. For each hit to 
%a block in the sampled set, the program counter (PC) of the relevant memory 
%instruction is hashed into the three tables and the corresponding counters 
%are decremented. For each eviction from a sampled set, the counters corresponding 
%to the PC of the last instruction to access the victim block are incremented. 
%For an LLC access, the predictor is consulted by hashing the PC of the memory 
%access instruction into the three tables and taking the sum of the indexed 
%counters. When the sum exceeds some threshold, the accessed block is 
%predicted to be dead. Tags in sampled sets are managed with true LRU and 
%a reduced associativity, but the LLC may be managed by any policy. 
%The paper applies its predictor to do block replacement and bypass for LLC.

Teran \textit{et al.} propose perceptron learning for reuse prediction for
bypass and replacement~\cite{Perc_Reuse}. Perceptron learning trains weights
selected by hashes of multiple features consisting of the PC of the memory
access instruction, some recent PCs, and two different shifts of the tag of
the referenced block. Features are hashed to index weights tables, reading out
weights that are and thresholded to make a prediction. When a block from one
of a few sampled sets~\cite{sdbp} is reused or evicted, corresponding weights
are decremented or incremented, respectively according to the perceptron
learning rule.  Multiperspective Reuse Prediction~\cite{Multiperspective}
improves on Perceptron by contributing many new features.

@@

\subsection{Base Prefetchers}

Two widely used prefetchers are Next-$n$ Lines~\cite{nextn} and
Stride~\cite{stride} prefetchers. Both capture regular memory access patterns
with low overhead. The Next-$n$ Lines prefetcher queues prefetches for the
next $n$ blocks after any given miss, expecting that the principle of spatial
locality will hold and those cache blocks after a missed block are likely to
be used in the future. The stride prefetcher identifies strided reference
patterns in programs based on past behavior of missing loads.  When a load
misses, cache blocks ahead of that miss are fetched in the pattern following
the previous behavior in the hope of avoiding future misses.  Without further
knowledge about temporal locality and application characteristics, these
prefetchers cannot do more than detecting and prefetching regular memory
access patterns with limited spatial locality.

Somogyi et al. proposed a practical, low-overhead prefetcher: Spatial Memory
Streaming (SMS) prefetcher~\cite{SMS}. SMS leverages code-based correlation to
take advantage of spatial locality in the applications over larger regions of
memory called spatial regions. It predicts future access pattern within a
spatial region around a miss based on a history of access patterns initiated
by that missing instruction in the past. SMS is effective, but it indirectly
infers future program control-flow when it speculates on the misses in a
spatial region. Thus, the state overhead of this predictor can be higher than
the others in this class. The idea was further extended by the class of
related prefetchers~\cite{STMS,Temporal_Instruction_Fetch}.

Other work also considers prefetch timeliness. Michaud proposed the
Best-Offset prefetcher~\cite{BOP}.  A prefetch offset is set automatically and
dynamically tries to adapt to the application behavior.  A good prefetch
offset is determined by looking at two recent access to a particular block
that did not happen too recently. The time between both accesses should be
greater than the latency for completing a prefetch request.

\subsection{Lookahead Prefetchers}
Previous prefetchers have only limited ability to capture memory delta
patterns. Moreover, a good prefetch proves to be useful only if the prefetcher
manages to bring the block before its next demand access. Both these issues
are overcome by Lookahead Prefetchers that use their own predictions to create
the speculative path of the program's memory access pattern. By going down the
speculation trail to a certain depth, the prefetcher can bring in cache blocks
well before they are demand accessed by the processor. VLDP~\cite{VLDP}
prefetches a static depth ahead of the demand fetch, without considering the
prefetching confidence. More recent Lookahead Prefetchers include SPP and
KPC~\cite{KPC}

@@

\subsection{Aggressive Prefetching}

@@

A low-accuracy aggressive prefetcher can significantly harm performance.  To
minimize interference from prefetching, Wu \textit{et. al} proposed
PACMan~\cite{pacman}, a prefetch-aware cache management policy.  PACMan
dedicates some LLC sets to each of three competing policies that treat demand
and prefetch requests differently, using the policy that shows the fewest
misses. Seshadri \textit{et. al.} proposed ICP~\cite{icp} that is also
designed as a comprehensive mechanism to mitigate prefetching interference. It
demotes a prefetch to the lowest priority on a demand hit based on the
observation that most prefetches are dead after their first hit.  To address
prefetcher-caused cache pollution, it also uses a variation of EAF~\cite{eaf}
to track prefetching accuracy and inserts only an accurate prefetch to the
higher priority position in the LRU stack.

\subsection{Machine Learning for Prefetching}
Peled \textit{et.al.} introduced interesting ideas for on-line Reinforcement
Learning and dynamically scaling the magnitude of feedback given to the
baseline prefetcher~\cite{Semantics}.  The prefetcher relies on compiler
support for getting features to build the context.  Liao \textit{et. al.}
focused on prefetching for data center applications~\cite{Datacenter}. They
used offline machine learning algorithms such as SVMs and logistic regression
to do a parametric search for an optimal prefetcher configuration.

>>>>

%Perceptron-based prediction mechanisms have proven to yield superior accuracy, 
%which represents an opportunity for optimizations in cache management. 
%Khan \textit{et. al.} proposed SDBP~\cite{sdbp}, in this work, a sampler 
%structure keeps partial tags of sampled sets separate from the cache. 
%Three tables of two-bit saturating counters are accessed using a technique 
%similar to a skewed branch predictor~\cite{Piece_Linear}. For each hit to 
%a block in the sampled set, the program counter (PC) of the relevant memory 
%instruction is hashed into the three tables and the corresponding counters 
%are decremented. For each eviction from a sampled set, the counters corresponding 
%to the PC of the last instruction to access the victim block are incremented. 
%For an LLC access, the predictor is consulted by hashing the PC of the memory 
%access instruction into the three tables and taking the sum of the indexed 
%counters. When the sum exceeds some threshold, the accessed block is 
%predicted to be dead. Tags in sampled sets are managed with true LRU and 
%a reduced associativity, but the LLC may be managed by any policy. 
%The paper applies its predictor to do block replacement and bypass for LLC.

In addition to branch prediction~\cite{PerceptronPredictor}, perceptron-based
learning has been applied to the area of cache management.
Teran \textit{et al.} propose using perceptrons to predict cache line resuse, 
bypass, and replacement~\cite{Perc_Reuse}. Perceptron Learning trains weights
selected by hashes of multiple features, including the PC of the memory
access instruction, some other recent PCs, and two different shifts of the tag of
the referenced block. These features are used to index into weight tables, and
the weights are then thresholded to generate a prediction. When a block from one
of a few sampled sets~\cite{sdbp} is reused or evicted, the corresponding weights
are decremented or incremented, according to the perceptron
learning rule.  Multiperspective Reuse Prediction~\cite{Multiperspective}
improves on Perceptron Learning by contributing many new features.

@@

\subsection{Spatial Prefetchers}

%Two widely used prefetchers are Next-$n$ Lines~\cite{nextn} and
%Stride~\cite{stride} prefetchers. Both capture regular memory access patterns
%with low overhead. The Next-$n$ Lines prefetcher queues prefetches for the
%next $n$ blocks after any given miss, expecting that the principle of spatial
%locality will hold and those cache blocks after a missed block are likely to
%be used in the future. The stride prefetcher identifies strided reference
%patterns in programs based on past behavior of missing loads.  When a load
%misses, cache blocks ahead of that miss are fetched in the pattern following
%the previous behavior in the hope of avoiding future misses.  Without further
%knowledge about temporal locality and application characteristics, these
%prefetchers cannot do more than detecting and prefetching regular memory
%access patterns with limited spatial locality.

Spatial prefetchers include such well-understood examples as the next-line (or
next-$n$-line) prefetcher, and the stream prefetcher, and are distinguished by prefetching
data without regard for the order in which the data will be accessed.
In addition to simpler these, Somogyi \textit{et al.} 
propose Spatial Memory Streaming (SMS)~\cite{SMS}.  SMS works by learning
the spatial footprint of all data used by a program within a region of memory,
and when that region is revisited by the program, the same spatial footprint
is prefetched.  Ishii \textit{et al.} propose the Access Map Pattern Matching 
prefetcher (AMPM)~\cite{AMPM}, which creates a map of all accessed lines within a region
of memory, and then analyzes this map on every access to see if any 
fixed-stride pattern can be identified and prefetched that is centered on 
the current access. Pugsley \textit{et al.} propose the Sandbox Prefetcher~\cite{sandbox},
which analyzes candidate fixed-offset prefetchers in a sandboxed environment
to determine which is most suitable for the current program phase.  Michaud
proposes the Best-Offset Prefetcher~\cite{BOP}, which determines the optimal
offset to prefetch while considering memory latency and prefetch timeliness.

%Somogyi et al. proposed a practical, low-overhead prefetcher: Spatial Memory
%Streaming (SMS) prefetcher~\cite{SMS}. SMS leverages code-based correlation to
%take advantage of spatial locality in the applications over larger regions of
%memory called spatial regions. It predicts future access pattern within a
%spatial region around a miss based on a history of access patterns initiated
%by that missing instruction in the past. SMS is effective, but it indirectly
%infers future program control-flow when it speculates on the misses in a
%spatial region. Thus, the state overhead of this predictor can be higher than
%the others in this class. The idea was further extended by the class of
%related prefetchers~\cite{STMS,Temporal_Instruction_Fetch}.

%Other work also considers prefetch timeliness. Michaud proposed the
%Best-Offset prefetcher~\cite{BOP}.  A prefetch offset is set automatically and
%dynamically tries to adapt to the application behavior.  A good prefetch
%offset is determined by looking at two recent access to a particular block
%that did not happen too recently. The time between both accesses should be
%greater than the latency for completing a prefetch request.

\subsection{Lookahead Prefetchers}

Unlike spatial prefetchers, lookahead prefetchers take program order into
account when they make predictions.  Shevgoor \textit{et al.} propose the Variable 
Length Delta Prefetcher (VLDP)~\cite{VLDP} that correlates histories of
deltas between cache line accesses within memory pages with the next delta
within that page.  SPP~\cite{SPP} and KPC's prefetching component~\cite{KPC} are
more recent examples of lookahead prefetchers.  They try to predict not only what
data will be used in the future, but also the precise order in which the data
will be used, within a given page.  Predictions made by lookahead prefetchers can
be fed back into their prediction mechanisms to predict even further down 
a speculative path of memory acesses. These prefetchers can also generalize their
learned patterns from one page, and use those patterns to make predictions in 
other pages.

%Previous prefetchers have only limited ability to capture memory delta
%patterns. Moreover, a good prefetch proves to be useful only if the prefetcher
%manages to bring the block before its next demand access. Both these issues
%are overcome by Lookahead Prefetchers that use their own predictions to create
%the speculative path of the program's memory access pattern. By going down the
%speculation trail to a certain depth, the prefetcher can bring in cache blocks
%well before they are demand accessed by the processor. VLDP~\cite{VLDP}
%prefetches a static depth ahead of the demand fetch, without considering the
%prefetching confidence. More recent Lookahead Prefetchers include SPP and
%KPC~\cite{KPC}

@@

\subsection{Managing Prefetched Data}

@@

A low-accuracy aggressive prefetcher can significantly harm performance.  To
minimize interference from prefetching, Wu \textit{et al.} propose
PACMan~\cite{pacman}, a prefetch-aware cache management policy.  PACMan
dedicates some LLC sets to each of three competing policies that treat demand
and prefetch requests differently, using the policy in the rest of the cache that shows the fewest
misses. Seshadri \textit{et al.} propose ICP~\cite{icp}, which
demotes a prefetch to the lowest reuse priority on a demand hit, based on the
observation that most prefetches are dead after their first hit.  To address
prefetcher-caused cache pollution, it also uses a variation of EAF~\cite{eaf}
to track prefetching accuracy, and inserts only accurate prefetches to the
higher priority position in the LRU stack.  Jain \textit{et al.} propose
Harmony~\cite{HarmonyHawkeye} to accomodate prefetches in their MIN 
algorithm-inspired Hawkeye cache management system.

\subsection{Machine Learning for Prefetching}

Peled \textit{et al.} introduce interesting ideas for on-line Reinforcement
Learning and dynamically scaling the magnitude of feedback given to the
baseline prefetcher~\cite{Semantics}.  The prefetcher relies on compiler
support for getting features to build the context.  Liao \textit{et al.}
focuse on prefetching for data center applications~\cite{Datacenter}. They
use offline machine learning algorithms such as SVMs and logistic regression
to do a parametric search for an optimal prefetcher configuration.
